---
title: "GeneralizedAdditiveModels"
author: "Anish Singh Walia"
date: "June 27, 2017"
output: html_document
---

# Generalized  Additive  Models


This is also a famous and very flexible technique of fitting and Modelling Non Linear Functions which are more flexible and fits data well.
In this technique we simply add __Non linear Functions__ on different variable to the Regression equation.

$$f(x) = y_i = \alpha \ +  \ f_1(x_1) \ + f_2(x_2) + \ .... + \ f_p(x_p)  $$

$$\text{where} \ f_p(x_p) \ \text {is a Non Linear function on} \  x_p \ variables.$$




```{r,message=FALSE,warning=FALSE , message=FALSE, warning=FALSE}
#requiring the Package 
require(gam)
require(ISLR)
attach(Wage)

gam1<-gam(wage~s(age,df=6)+s(year,df=6)+education ,data = Wage)
#in the above function s() is the shorthand for fitting smoothing splines in gam() function
summary(gam1)
#Plotting the Model
par(mfrow=c(1,3))
plot(gam1,se = TRUE)



```
In the above Plots the Y-axis contains the Non Linear functions and x-axis contains the Predictors used in the Model.The Whole Model is *__Additive__* in nature.


---


### We can also fit a Logistic Regression Model using gam()

```{r}
#logistic Regression Model
gam2<-gam(I(wage >250) ~ s(age,df=4) + s(year,df=4) +education , data=Wage,family=binomial)



```


