---
title: "R Notebook"
output: html_notebook
---

#Spilnes

Splines are used to add __Non linearities__ to a Linear Model and it is a Flexible Technique than Polynomial Regression.Splines Fit the data very *__smoothly__* in most of the cases where
polynomials would become wiggly and overfit the training data.




```{r,message=FALSE,warning=FALSE}
#loading the Splines Packages
require(splines)
#ISLR contains the Dataset
require(ISLR)
attach(Wage)

```

---


### Fitting a Cubic Spline with 3 Knots(Cutpoints)
What It does is that it transforms the Regression Equation by transforming the Variables with a truncated *__Basis__* Function- $$b(x)$$ ,with continious derivatives upto order 2.

The Regression Equation Becomes -

$$y_i = \alpha + \beta_1.b_1(x_i)\ +  \beta_2.b_2(x_i)\ + \ .... \beta_{k+3}.b_{k+3}(x_i) \ + \epsilon_i   $$

where $$\bf b_n(x_i)\ is \  The \  Basis \ Function$$.
```{r}
#3 cutpoints at ages 25 ,50 ,60
fit<-lm(wage ~ bs(age,knots = c(25,40,60)),data = Wage )
summary(fit)


```

