---
title: "Splines"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---

#What are Splines ?

Splines are used to add __Non linearities__ to a Linear Model and it is a Flexible Technique than Polynomial Regression.Splines Fit the data very *__smoothly__* in most of the cases where polynomials would become wiggly and overfit the training data.Polynomials 
tends to get wiggly and fluctuating at the tails which sometimes __overfits__ the training data and doesn't *__generalizes well due to high variance at high degrees of polynomial__.*



```{r,message=FALSE,warning=FALSE}
#loading the Splines Packages
require(splines)
#ISLR contains the Dataset
require(ISLR)
attach(Wage)

agelims<-range(age)
#Generating Test Data
age.grid<-seq(from=agelims[1], to = agelims[2])

```

---


### Fitting a Cubic Spline with 3 Knots(Cutpoints)
What It does is that it transforms the Regression Equation by transforming the Variables with a truncated *__Basis__* Function- $$b(x)$$ ,with continious derivatives upto order 2.

$$\textbf{The order of the continuity}= (d - 1) , \ where \ d \ is \ the \ number \ of \ degrees \ of \ polynomial$$

$$\textbf {The Regression Equation Becomes}$$ -

$$f(x) = y_i = \alpha + \beta_1.b_1(x_i)\ +  \beta_2.b_2(x_i)\ + \ .... \beta_{k+3}.b_{k+3}(x_i) \ + \epsilon_i   $$

 $$\bf where \ \bf b_n(x_i)\ is \  The \  Basis \ Function$$.
 
$$\text{The idea here is to transform the variables and add a linear combination of the variables using the Basis power function to the regression function f(x).} $$
```{r}
#3 cutpoints at ages 25 ,50 ,60
fit<-lm(wage ~ bs(age,knots = c(25,40,60)),data = Wage )
summary(fit)
#Plotting the Regression Line to the scatterplot   
plot(age,wage,col="grey",xlab="Age",ylab="Wages")
points(age.grid,predict(fit,newdata = list(age=age.grid)),col="darkgreen",lwd=2,type="l")
#adding cutpoints
abline(v=c(25,40,60),lty=2,col="darkgreen")

```
The above Plot shows the smoothing and local effect of Cubic Splines , whereas Polynomias might become wiggly and the tail.*__The cubic splines have continious 1st Derivative and continious 2nd derivative.__*

--------


###Smoothing Splines 

These are mathematically more challenging but they are more smoother and flexible as well.It does not require the selection of the number of Knots , but require selection of only a __Roughness Penalty__ which accounts for the wiggliness(fluctuations) and controls the roughness of the function and variance of the Model.

$$ \text{Let the RSS(Residual Sum of Squares) be} \ { g(x_i) }$$
$$ \ minimize \ { g \in RSS} :\ \sum\limits_{i=1}^n ( \ y_i \ - \ g(x_i) \ )^2 + \lambda \ \int g''(t)^2 dt , \quad  \lambda > 0$$
$$ \text{ where },\ \lambda \int g''(t)^2 dt \  {is \ called \ the \ Roughness \ Penalty. }  $$

$$ \textbf {The Roughness Penalty controls how wiggly g(x) is. The smaller the }  \lambda , \textbf{ the more wiggly and  fluctuating the function is. }$$  

$$\textbf {As} \  \lambda ,\textbf {approcahes}\ \infty , \textbf {the function} \ g(x) \textbf  { becomes linear} . $$

In smoothing Splines we have a __Knot__ at every unique value of $x_i$ .


```{r}
fit1<-smooth.spline(age,wage,df=16)
plot(age,wage,col="grey",xlab="Age",ylab="Wages")
points(age.grid,predict(fit,newdata = list(age=age.grid)),col="darkgreen",lwd=2,type="l")
#adding cutpoints
abline(v=c(25,40,60),lty=2,col="darkgreen")
lines(fit1,col="red",lwd=2)
legend("topright",c("Smoothing Spline with 16 df","Cubic Spline"),col=c("red","darkgreen"),lwd=2)

```

---


### Implementing Cross Validation to select value of $\lambda$  and Implement Smoothing Splines 


```{r}
fit2<-smooth.spline(age,wage,cv = TRUE)
fit2
#It selects $\lambda=6.794596$ it is a Heuristic and can take various values for how rough the function is

plot(age,wage,col="grey")
#Plotting Regression Line
lines(fit2,lwd=2,col="purple")
legend("topright",("Smoothing Splines with 6.78 df selected by CV"),col="purple",lwd=2)


```

####This Model is also very Smooth and Fits the data well. 

---



